Supporting Information: Unsupervised embedding of trajectories captures the latent structure of mobility

### 2022-03-15 paper/current/supp.tex

% S11 TEXT
\paragraph*{S12 Text}
\label{si:text:disambiguation}
{\bf Derivation of Eq.\ref{eq:nce} - Noise Contrastive Estimation}

@@

@@

>>>>

% S12 TEXT
\paragraph*{S12 Text}
\label{si:text:disambiguation}
{\bf Derivation of Eq.\ref{eq:nce} - Noise Contrastive Estimation}

@@

%
% S15 TEXT
\paragraph*{S15 Text}
\label{si:text:generalized_gravity_model}
{\bf Generalized gravity model.}

Generalized gravity models use external information about each location to inform mobility predictions. 
    Here, we test whether a generalized gravity model that includes information about shared city, region, nation, continent, and language alongside geographic distance offers better predictions than embedding distance. 
    We specify the generalized model as follows,
    \begin{equation}
      \log{\frac{T_{ij}}{M_iM_j}} =  \gamma \log{d_{ij}} +  \sum_{x} \alpha^{x}*{g_{ij}^x} +  \sum_{x} \beta^{x}*{l_{ij}^x} 
   \end{equation}
    where $g_{ij}^{x}, x \in [\text{city, region, nation, and continent}]$ is a dummy variable that is given a value of 1 when two organizations are located in the same geographic city, region, nation, or continent, respectively; 
    otherwise, dummy variable is zero.
    Similarly, $l_{ij}^x,  x \in [\text{language, language family}]$ is an dummy variable that is set to 1 if two organizations are in countries that share the same language or language family, respectively.

    Better results are obtained using the generalized gravity model ($R^2=0.242$, $RSME=0.861$, and $CPC=0.383$) compared to standard gravity model using only geographical distance ($R^2=0.219$, $RSME=0.874$, and $CPC=0.373$).
    However, the gravity model using embedding distance results in the best performance by a wide margin ($R^2=0.482$, $RSME=0.711$, and $CPC=0.459$). 
    These findings demonstrate how the embedding space can encode the complex and multi-faceted nature of mobility into a single robust vector-space representation, even without any external information provided.

@@

%
% Figure — CPC Measures
% 
\begin{figure}
    \centering
    \includegraphics{\figdir/Descriptive/sizeall_precedence_exp_global_cpc-measures_d300_ws1_gamma1.0.pdf}
    \caption{\textbf{Embedding distance outperforms alternatives in common part of commuters (CPC).}
    The CPC measure is calculated for both the power and exponential form of the gravity model, with the highest value reported for each distance type. 
    }
    \label{fig:supp:cpc-performance}
\end{figure}
